Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2018-09-13 RNARedPrint.tex
Update P(S|R)~P(R|S) corelation plots in supplement

We observe that
targeting realistic RNA energies in the realistic
Turner RNA energy model
works well by performing sampling based on less complex RNA energy models, which cause low treewidth. This result is highly relevant in practice, since it allows to combine high efficiency with high accuracy and moreover leaves room for extensions of the method.

Eventually this strategy allows generating biologically relevant
multi-target designs, as we demonstrate in our application 
to a large set of multi-target RNA design instances from a
representative benchmark.

>>>>

We observe that
targeting realistic RNA energies in the realistic
Turner RNA energy model
works well by performing sampling based on less complex RNA energy models, which keeps the treewidth low. This result is highly relevant in practice, since it allows to combine high efficiency with high accuracy and moreover leaves room for extensions of the method.

Eventually this strategy allows generating biologically relevant
multi-target designs, as we demonstrate in our application 
to a large set of multi-target RNA design instances from a
meaningful, representative benchmark.

### 2018-09-13 supplement.tex
Update P(S|R)~P(R|S) corelation plots in supplement

Already previous work on RNA Design, e.g. INFO-RNA and Incarnation, observed
that---briefly put---negative design benefits from positive design. Here, more
concretely, we only \emph{demonstrate} this effect (for small examples), where
we consider \Def{negative design} as  
designing RNA sequences that fold with high probability $Pr(R|S)$ into the
given target structure $R$; and \Def{positive design}, as generating one or
many sequences that have a low energy for $R$.

For a double-stem structure, we produce a uniform and a weighted Boltzmann
sampling, each of 10000 sequenes. For each sampled sequence $S$, we determine
the Turner energy $E(S,R)$ as well as the
%difference $E_\Delta(S,R)$ of $E(S,R)$ to the ensemble energy $E_{\text{ens}}(S)$ of $S$ 
probability $Pr(R|S)$ of the structure $R$ in the ensemble of sequence $S$ 
(where the energies are calculated in the Turner model using RNAfold).

Note that low energy $E(S|R)$ is (non-linearly) correlated with high
probability $Pr(S|R)$, the probability to see $S$ in the Boltzmann distribution
of sequences folding into $R$. Recall that the probability $Pr(R|S)$ can be
calculated as $\exp(-E(S,R)/RT)/Z(S)$,
where $Z(S)=\sum_{R'} \exp(-E(S,R')/RT)$. 
%
Consequently, Figure~\ref{fig:negforpos} ultimately illustrates the dependency
between $Pr(S|R)$ and $Pr(R|S)$ for uniform samples and low energy samples
(where the latter can be efficiently generated due to our Boltzmann sampling
strategy).

@@

\begin{figure}
\textbf{A}~\includegraphicstop[width=0.48\textwidth]{Figs/Plots/doublestem_weight1}%
\textbf{B}~\includegraphicstop[width=0.48\textwidth]{Figs/Plots/doublestem_weight500}
\caption{Comparision of (A) Uniform sampling to (B) Boltzmann sampling with
weight $\pi=500$. For given structure $R$, we plot the energies $E(S,R)$ of
sampled sequences against the probabilities $Pr(R|S)$; the former is controlled
by Boltzmann sampling, while the latter is a common negative design objective.}
\label{fig:negforpos}
\end{figure}

>>>>

Already previous work on RNA Design, e.g. INFO-RNA and Incarnation, observed
that---put briefly---negative design benefits from positive design. Here, more
concretely, we simply \emph{demonstrate} this effect (for small examples), where
we consider \Def{negative design} as  
designing RNA sequences that fold with high probability $Pr(R|S)$ into the
given target structure $R$; and \Def{positive design}, as generating one or
many sequences that have a low energy for $R$.

For each of two small examples, a single-stem and a double-stem structure, we produce a uniform and a weighted Boltzmann
sampling, each of 10000 sequenes. For each sampled sequence $S$, we determine
the Turner energy $E(S,R)$ as well as the
%difference $E_\Delta(S,R)$ of $E(S,R)$ to the ensemble energy $E_{\text{ens}}(S)$ of $S$ 
probability $Pr(R|S)$ of the structure $R$ in the ensemble of sequence $S$ 
(where the energies are calculated in the Turner model using RNAfold).

Note that low energy $E(S|R)$ is (non-linearly) correlated with high
probability $Pr(S|R)$, the probability to see $S$ in the Boltzmann distribution
of sequences folding into $R$. Recall that the probability $Pr(R|S)$ can be
calculated as $\exp(-E(S,R)/RT)/Z(S)$,
where $Z(S)=\sum_{R'} \exp(-E(S,R')/RT)$.
%
Consequently, Figure~\ref{fig:negforpos} ultimately illustrates the dependencies
between $Pr(S|R)$ and $Pr(R|S)$ for uniform samples and low energy samples
(where those samples can be efficiently generated due to our Boltzmann sampling
strategy). For our examples, we observe (Fig.~\ref{fig:negforpos}) that Boltzmann sampling with weight 500 (green), generates good negative designs 
with substantially higher probability than uniform sampling (blue).

@@

\begin{figure}
\textbf{A}~\includegraphicstop[width=0.48\textwidth]{Figs/Plots/stem_weight1-500}%
\textbf{B}~\includegraphicstop[width=0.48\textwidth]{Figs/Plots/doublestem_weight1-500}%
\caption{Comparision of uniform sampling and Boltzmann sampling with
weight $\pi=500$. We show results for two different small structures: (A) single stem structure (B) double stem structure. For each respective structure $R$, we plot the energies $E(S,R)$ of
sampled sequences against the probabilities $Pr(R|S)$; the former is controlled
by Boltzmann sampling, while the latter is a common negative design objective.}
\label{fig:negforpos}
\end{figure}
